\specialchapt{ACKNOWLEDGEMENTS}

First, I would like to thank my wife, Rianka, for her constant support, patience, understanding and encouragement throughout my graduate education. I would also like to thank my younger brother, Upamanyu, who have always supported and stood by me in my low moments.

I would like to thank my major professor, Dr. Dianne Cook for her valuable guidance and support. The various statistical methods learnt from her, specially on data visualizations and academic writing, will help me excel in my future endeavors. But her sense of humor, calmness and balance in life have had the biggest effects on me and it would help me to improve as a human being.

The suggestions provided by Dr. Heike Hofmann have helped improving the quality and scope of my research work. I am very thankful to her.

I really appreciate the valuable comments and suggestions made by my committee members. I would also like to thank them for their time and patience. 

Graphics Working Group has provided me various suggestions and participated in a number of pilot studies. I would like to thank everyone in the group. I would specially thank Mahbubul Majumder for helping me whenever I needed.

I would also like to thank the Department of Statistics in Iowa State University for providing me this opportunity to pursue my doctorate degree. The inspiring teaching styles of many professors have helped me understand and learn the subject. 

Finally, I thank my family back home and all my friends in Ames who have been more than my family and supported me unconditionally during my graduate education.
